---
title: "Final Project"
author: "Kenan"
date: "12/24/2020"
output: html_document
---

# Storm track dataset

Description
This data is a subset of the NOAA Atlantic hurricane database best track data, https://www.nhc.noaa.gov/data/#hurdat. The data includes the positions and attributes of 198 tropical storms, measured every six hours during the lifetime of a storm.

* Format - A tibble with 10,010 observations and 13 variables:
* name - Storm Name
* year,month,day - Date of report
* hour - Hour of report (in UTC)
* lat,long - Location of storm center
* status - Storm classification (Tropical Depression, Tropical Storm, or Hurricane)
* category - Saffir-Simpson storm category (estimated from wind speed. -1 = Tropical Depression, 0 = Tropical Storm)
* wind - storm's maximum sustained wind speed (in knots)
* pressure - Air pressure at the storm's center (in millibars)
* ts_diameter - Diameter of the area experiencing tropical storm strength winds (34 knots or above)
* hu_diameter - Diameter of the area experiencing hurricane strength winds (64 knots or above)
"

```{r}
library("ggplot2")
library('dplyr')
library("wordcloud")
library("reshape2")
options(warn=-1)
```

> Read in the dataset and remove unnessary columns
```{r}
df <- read.csv('https://vincentarelbundock.github.io/Rdatasets/csv/dplyr/storms.csv')
df$date <- paste(df$year, df$month, df$day, sep='-')
df <- df[ , !(names(df) %in% c('month', 'day', 'hour', 'X'))]
```

Global statistics
```{r}
summary(df)
```

> From 1975 to 2015 how many of storms have occurred each year
```{r}
ydf <- as.data.frame(table(df$year))
colnames(ydf) <- c('Year', 'Frequency')
ggplot(ydf, aes(x = Year, y = Frequency)) + geom_col(aes(fill = Frequency)) + geom_text(aes(label = Frequency), angle = 90, nudge_y=30) + scale_x_discrete(guide = guide_axis(angle = 90))
```

> How do the storms vary by category for each year
```{r}
tb1 <- table(df$status, df$year)
barplot(tb1, xlab='Year', ylab='Status Counts', col=c('red','green','blue'))
legend(x=1, y=575, legend=unique(df$status), fill=c('red','green','blue'))
```

> What makes a storm change status
```{r}
qplot(df$wind, df$pressure, main='Pressure vs Wind', xlab='Wind (knots)', ylab='Pressure (millibars)', col=df$status) + theme(plot.title = element_text(hjust = 0.5))
```

> Which storm name was most popular?
```{r}
cdf <- count(df, name)
wordcloud(words = cdf$name, freq = cdf$n, color = 'blue', size = 1, shape = "rectangle", backgroundColor = "white") 
```

> Which storm did the most damage
```{r}
hdf <- df %>% group_by(name) %>% summarise(Hurricane = sum(hu_diameter), Tropical_Storm = sum(ts_diameter))
hdf <- filter(hdf, Tropical_Storm > 0)
hdf <- melt(hdf, id.vars ="name", measure.vars = c("Hurricane", "Tropical_Storm"))
colnames(hdf) <- c('name', 'Storm_Type', 'value')

ggplot(hdf, aes(name, value, colour = Storm_Type)) + labs(title="Storm Damage", x ="Name", y = "Area (miles^2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(angle = 90)) + geom_point()
```

> Most recent storm paths
```{r}
rdf <- filter(df, year == 2015)
head(rdf)
#geoplot(lat=df$lat, lon=df$long)
```

