---
title: "KSooklall_Homework11"
author: "Kenan Sooklall"
date: "4/8/2021"
output:
  pdf_document: default
  html_document: default
---

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r cars}
summary(cars)
```

```{r}
lmodel = lm(dist ~ speed, data=cars)
summary(lmodel)
```

For every unit of speed (mph) that a car travels the distance the car will stop when the breaks is slammed is 4ft. The $r^2$ of 0.65 implies that 65% of the variance on the distance the car traveled is explained by the speed.

```{r}
cars %>% ggplot(aes(speed, dist)) + geom_point() + stat_smooth(method = "lm")
```

```{r}
plot(fitted(lmodel),resid(lmodel))
```

The residuals are some what scattered above and below zero. Overall, this plot tells us that using the regression model is an ok predictor between speed and distance.