---
output:
  pdf_document: default
  html_document: default
---
## Problem 1


(1) Show that $A^T A \neq A A^T$ in general. (Proof and demonstration.)
(2) For a special type of square matrix A, we get A T A = AA T . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

### Proof by contradiction

#### Let: 

  $A$ be a matrix of shape $m x n$ where $m$ and $n$ are in $\mathbb{N}$ and $m \neq n$

#### Suppose:

  $A^T A = A A^T$

#### Then:

  $n x m \times m x n = m x n \times n x m$ ->  $n x n = m x m$

That equation is only true when $n=m$ which contradicts our assumption that $m\neq n$

#### Therefore:

$A^T A \neq A A^T$

#### Exception:

$m=n$ then $A$ is a square matrix and consequently $A^T A = A A^T$ will hold 


# Problem 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. Please submit your response in an R Markdown document using our class naming
convention, E.g. LFulton_Assignment2_PS2.png

You don’t have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire
assignment in R, then please submit only one markdown document for both the problems.

```{r}
factorize <- function(X){
  n <- NROW(X)
  U <- X
  L <- diag(n)
  
  for (j in c(1:n)){
    for(i in c(2:n)){
      if(i > j){
        r <- U[j, ]
        v <- U[i, j] / r[j]
        U[i,] <- U[i,] - (v * r)
        L[i, j] <- v
      }
    }
  }
  
  return (list(L=L, U=U))
}
```

```{r}
A <- matrix(c(1,2,3,1,1,1,2,0,1), nrow=3)
f <- factorize(A)
B <- f$L %*% f$U

sum(B == A) == length(A)
```