---
title: "Homework 2"
author: "Kenan Sooklall"
date: "12/20/2020"
output: html_document
---
```{r}
library(RCurl)
```
# Dataset - Fertility and Contraception

Data - https://vincentarelbundock.github.io/Rdatasets/csv/carData/Robey.csv

Docs - https://vincentarelbundock.github.io/Rdatasets/doc/carData/Robey.html

Column data:

* region -A factor with levels: Africa; Asia, Asia and Pacific; Latin.Amer, Latin America and Caribbean; Near.East, Near East and North Africa.
* tfr - Total fertility rate (children per woman).
* contraceptors - Percent of contraceptors among married women of childbearing age.

```{r}
df <- read.csv('https://vincentarelbundock.github.io/Rdatasets/csv/carData/Robey.csv')
```

#### 1. Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes.
```{r}
summary(df)
```
World population statistics
```{r}
print(mean(df$tfr))
print(median(df$contraceptors))
```
#### 2. Create a new data frame with a subset of the columns and rows. Make sure to rename it.
```{r}
sdf <- subset(df, region == 'Latin.Amer', select=c(region, tfr, contraceptors))
sdf
```
This subset trims the dataframe from 50 rows 4 columns to 16 rows and 3 columns
#### 3. Create new column names for the new data frame.
```{r}
colnames(sdf) <- c('region', 'tfr', 'contra')
```
I shortened contraceptors to contra

#### 4. Use the summary function to create an overview of your new data frame. The print the mean and median for the same two attributes. Please compare.
```{r}
summary(sdf)
```
For just Latin America the mean is smaller but the median is higher
```{r}
print(mean(sdf$tfr))
print(median(sdf$contra))
```
#### 5. For at least 3 values in a column please rename so that every value in that column is renamed. For example, suppose I have 20 values of the letter “e” in one column. Rename those values so that all 20 would show as “excellent”.
```{r}
df$region <- as.character(df$region)
df$region[df$region == 'Latin.Amer'] <- "LatinAmerica"
df$region[df$region == 'Near.East'] <- "East"
df$region[df$region == 'Africa'] <- "africa"
```

#### 6. Display enough rows to see examples of all of steps 1-5 above.
Subset df filtered for LatinAmerica and column values renamed
```{r}
print(sdf)
```
Original df with regions renamed based on question 5
```{r}
print(df)
```

#### 7. BONUS – place the original .csv in a github file and have R read from the link. This will be a very useful skill as you progress in your data science education and career.
On github locate the csv file and click raw on the top right then use that link
```{r}
link <- 'https://raw.githubusercontent.com/ksooklall/CUNY-SPS-Masters-DS/main/bridge_r/w2/fertility_and_contraception.csv'
text_data <- getURL(link)
github_df <- read.csv(text=text_data)
head(github_df)
```