---
title: "Homework 1"
author: "Kenan Sooklall"
date: '2022-01-28'
output:
  html_document: default
  pdf_document: default
---

```{r, echo=F, warning=F, message=F}
library(tidyverse)
```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
df <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r}
head(df)
```
```{r}
summary(df)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Top 10 growing industries, city and states by revenue

```{r}
df %>% group_by(Industry) %>% summarise(total=sum(Revenue)) %>% arrange(desc(total)) %>% top_n(total, n=10)
df %>% group_by(City) %>% summarise(total=sum(Revenue)) %>% arrange(desc(total)) %>% top_n(total, n=10)
df %>% group_by(State) %>% summarise(total=sum(Revenue)) %>% arrange(desc(total)) %>% top_n(total, n=10)
```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r fig.width=7, fig.height=7}
df %>% 
  group_by(State) %>% 
  summarise(counts=n_distinct(Name)) %>% 
  ggplot(aes(x=reorder(State, counts), y=counts)) + 
    geom_col() + 
    coord_flip() + 
    labs(x='States', y='Number of Companies', title='Number of companies by states')
```

California has the largest number of companies followed by Texas and New York. There are very few companies in wyoming, West virginia, Arkansas, Puerto Rico.

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
df %>% 
    filter(complete.cases(Employees)) %>% 
    filter(State == 'NY') %>% 
    group_by(Industry) %>% 
    summarise(Employee_averages=mean(Employees), Employee_median=median(Employees)) %>% 
  gather(key = key, value = value, Employee_averages:Employee_median) %>% 
  ggplot(aes(x=Industry, y=value)) +
    geom_boxplot() + 
    coord_flip() + 
    labs(title='NY employment by industries', y='Employed')
```

The state with the 3rd most companies is NY. In NY the Business Products & Services industry has the largest range and the most number of employees employed. Customer Products & Services follow next then Travel & Hospitality and Human Resources. All other industries have a very small range between their mean and median number of employees.

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
rev_by_industry <- df %>% 
  group_by(Industry) %>% 
  summarise(rev=sum(Revenue))

rev_by_industry %>% 
  ggplot(aes(x=reorder(Industry,rev), y=rev)) + 
    geom_col() + 
    coord_flip() + 
    labs(x='Industries', y='Revenue', title='Revenue by Industries')

rev_by_industry %>% arrange(desc(rev))
```

The business products & services industry generates the most revenue at 26 billion followed by IT services at 20 billion.