---
title: "KSooklall Homework 12 DATA 605"
author: "Kenan Sooklall"
date: "4/19/2021"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(ggplot2)
```

who.csv dataset contains real-world data from 2008.

Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
TotExp: sum of personal and government expenditures.

```{r}
path = '/home/kenan/Documents/learning/masters/CUNY-SPS-Masters-DS/DATA_605/homeworks/homework12/'
df = read.csv(paste0(path, 'who.csv'))
```

### Exercise 1 
Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss
whether the assumptions of simple linear regression met.

```{r}
df %>% ggplot(aes(x=LifeExp, y=TotExp)) + geom_point() +   stat_smooth(method = "lm")

model <- lm(LifeExp~TotExp, data=df)
summary(model)
```

An $R^2$ of 0.2577 is very low and states there isn't a strong correlation between LifeExp and TotExp.
The standard error is 7.795e-06 and the p value is very small 7.71e-14.
F-statistic = 65.2  which is the ratio of two variances (SSR/SSE), the variance explained by the parameters in the model (sum of squares of regression, SSR) and the residual or unexplained variance (sum of squares of error, SSE).

### Exercise 2
Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06
power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r
re-run the simple regression model using the transformed variables. Provide and interpret the F
statistics, R^2, standard error, and p-values. Which model is "better?"

```{r}
df$lifexp46 <- df$LifeExp ^ (4.6)
df$totexp06 <- df$TotExp ^ (0.06)

df %>% ggplot(aes(x=lifexp46, y=totexp06)) + geom_point() +   stat_smooth(method = "lm")

model <- lm(lifexp46~totexp06, data=df)
summary(model)
```

Very surprising results. The R^2 tripled making the relationship look more linear. The F-statistic and the std error also went up as well with the P-vaues about the same ~0. This model is "better" than the last.

### Exercise 3
Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life
expectancy when TotExp^.06=2.5.

```{r}
totalexp06 = c(1.5, 2.5)
yhat1 = predict(model, data.frame(totexp06=totalexp06[1]))
print(yhat1)

yhat2 = predict(model, data.frame(totexp06=totalexp06[2]))
print(yhat2)
```

### Exercise 4
Build the following multiple regression model and interpret the F Statistics, R^2, standard error,
and p-values. How good is the model?

$LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp$

```{r}
model = lm(LifeExp ~ PropMD + TotExp + (PropMD * TotExp), data=df)
summary(model)
```

This model is more complex since there are more variables. The statistic values are better than the model from exercise 2 but worst than exercise 3.

### Exercise 5
Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why
or why not?

```{r}
print(predict(model, data.frame(PropMD=0.03, TotExp=14)))

df %>% ggplot(aes(LifeExp)) + geom_histogram()
```

The forecast is realistic, many have lived to 108 years old; however that is very far above the average in the real world and would be an extrapolation value in this data set.