---
title: "KSooklall_Homework10 DATA 607"
author: "Kenan Sooklall"
date: "4/16/2021"
output:
  pdf_document: default
  html_document: default
---

The code in this report are taken from the textbook:
Text Mining with R
Chapter 2 - Sentiment analysis with tidy data

```{r}
library(tidytext)
library(tidyverse)
library(ggplot2)
library(wordcloud)
```

Load sentiment data sets

```{r}
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

Read in the data and plot the ratio of good vs bad sentiment
```{r}
df = read.csv('https://raw.githubusercontent.com/ksooklall/CUNY-SPS-Masters-DS/main/DATA_607/homework/homework10/imbd_movie_reviews.data')
df %>% ggplot(aes(x=sentiment)) + geom_bar()
```

Unnest the review column and remove stop_words.
Stop_words are words that carry no sentiment like [i, the, a, able, about ...]
```{r}
text <- df %>% mutate(linenum = row_number()) %>% unnest_tokens(word, review) %>% anti_join(stop_words)
```

View the distribution of different sentiment words
```{r}
text %>% select(word) %>%
    inner_join(get_sentiments("bing")) %>%
    count(word, sort = TRUE) %>% top_n(n, n=10) %>% ggplot(aes(x=word, y=n)) + geom_col() + coord_flip()
```

```{r}
text %>% select(word) %>%
    inner_join(get_sentiments("nrc")) %>%
    count(word, sort = TRUE) %>% top_n(n, n=10) %>% ggplot(aes(x=word, y=n)) + geom_col() + coord_flip()
```

```{r}
text %>% select(word) %>%
    inner_join(get_sentiments("afinn")) %>%
    count(word, sort = TRUE) %>% top_n(n, n=10) %>% ggplot(aes(x=word, y=n)) + geom_col() + coord_flip()
```


Word clouds

```{r}
text %>% pull(word) %>% wordcloud(min.freq = 10, max.word=100)
```


```{r}
library(reshape2)

text %>% select(word) %>% inner_join(get_sentiments("afinn")) %>%inner_join(get_sentiments("nrc")) %>%inner_join(get_sentiments("bing")) %>% count(word, sentiment, sort=T) %>% acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("red", "blue"),
                   max.words = 100)
```