---
title: "Ksooklall_Homework4"
author: "Kenan Sooklall"
date: "2/10/2021"
output: html_document
---

Let 

\[
A = 
  \begin{bmatrix}
    1 & 2 & 3   \\
    -1 & 0 & 4  \\
  \end{bmatrix}
\]

compute $X = AA^T$ and $Y = A^TA$

```{r}
A <- matrix(c(c(1,-1), c(2, 0), c(3,4)), nrow=2, ncol=3)
At <- t(A)
```

```{r}
X = A %*% At
X
```

```{r}
Y = At %*% A
Y
```
Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R.

```{r}
Xe <- eigen(X)
print(Xe$values)
print(Xe$vectors)
```

```{r}
Ye <- eigen(Y)
print(Ye$values)
print(Ye$vectors)
```
Compute the left-singular, singular values, and right-singular vectors of $A$

```{r}
s <- svd(A)
```

Left singular

```{r}
ls <- s$u
```

Right singular

```{r}
rs <- s$v
```

Singular

```{r}
singular <- s$d
```

Check left and right singular with eigen vectors

```{r}
print(all.equal(abs(Xe$vectors), abs(ls)))
print(all.equal(abs(Ye$vectors), abs(rs)))
```

In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the
non-zero singular values of A.

```{r}
round(Xe$values, 3)
round(Ye$values, 3)
```

Indeed the first two values are the same and the third of Ye is 0

# Load images

Set up global variables

```{r}
library(EBImage)
library(OpenImageR)
library(jpeg)

path <- "/home/kenan/Documents/learning/masters/CUNY-SPS-Masters-DS/DATA_605/homeworks/homework4/images/"
n=17
files=list.files(path, pattern="*.jpg")[1:n]
height=1200
width=2500
scale=20
```

Function to read in images from path and plot them


```{r}
plot_image = function(img)
{ 
  #res = dim(img)[2:1] # get the resolution, [x, y]
  #plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]),asp=1,type='n',xaxs='i',yaxs='i',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
  imageShow(img)
}
```

resize and populate vector $im $ with all images $dim(im) = (17, 60, 125, 3)

```{r}
im=array(rep(0,length(files)*height/scale*width/scale*3), dim=c(length(files), width/scale, height/scale, 3))

for (i in 1:n){
  im[i,,,]=resizeImage(readImage(paste0(path, files[i])),width=width/scale, height=height/scale)
  }
```

Plot

```{r}
par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))

for (i in 1:9){  #plot the first images only
  plot_image(im[i,,,])
}
```

Flatten each vector so dim(im) is 17 x 22500

```{r}
#flat=matrix(0, n, prod(dim(im)))
flat=matrix(0, n, prod(dim(im)[2:4]))
for (i in 1:n) {
  newim <- readImage(paste0(path, files[i]))
  r=as.vector(im[i,,,1]) 
  g=as.vector(im[i,,,2])
  b=as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
shoes=as.data.frame(t(flat))
```

Scale all values to have mean of 0 and standard deviation of 1

```{r}
scaled=scale(shoes, center = TRUE, scale = TRUE)
mean.shoe=attr(scaled, "scaled:center") #saving for classification
std.shoe=attr(scaled, "scaled:scale") 
Sigma_=cor(scaled)
myeigen=eigen(Sigma_)
cumsum(myeigen$values) / sum(myeigen$values)
```

```{r}
scaling=diag(myeigen$values[1:5]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes=scaled%*%myeigen$vectors[,1:5]%*%scaling
imageShow(array(eigenshoes[,1], c(125,60,3)))
```


```{r}
height=1200
width=2500
scale=20
newdata=im
dim(newdata)=c(length(files),height*width*3/scale^2)
mypca=princomp(t(as.matrix(newdata)), scores=TRUE, cor=TRUE)
```

```{r}
mypca2=t(mypca$scores)
dim(mypca2)=c(length(files),height/scale,width/scale,3)
par(mfrow=c(5,5))
par(mai=c(.001,.001,.001,.001))
for (i in 1:n){#plot the first 25 Eigenshoes only
plot_image(mypca2[i,,,], from_file = FALSE)  #complete without reduction
}
```

```{r}
a=round(mypca$sdev[1:n]^2/ sum(mypca$sdev^2),3)
cumsum(a)
```

