---
title: "Principal Components Analysis (PCA)"
author: "Kenan Sooklall"
date: "2/10/2021"
output:
  pdf_document: default
  html_document: default
---



The goal of this homework is to show how PCA can take advantage of the linear combination between pixel values to create a lower dimensional representation.

Set up global variables

```{r}
library(EBImage)
library(OpenImageR)
library(jpeg)

path <- "/home/kenan/Documents/learning/masters/CUNY-SPS-Masters-DS/DATA_605/homeworks/homework4/images/shoes/"

files=list.files(path, pattern="*.jpg")
n=length(files)

# shoes dim
height=1200
width=2500
scale=20

# Pokemon dim
#height=224
#width=224
#scale=2
```

Function to read in images from path and plot them


```{r}
plot_jpeg = function(path, add=FALSE)
{ jpg = readJPEG(path, native=T) # read the file
  res = dim(jpg)[2:1] # get the resolution, [x, y]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]),asp=1,type='n',xaxs='i',yaxs='i',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
  rasterImage(jpg,1,1,res[1],res[2])
}
```

Resize each image and populate vector $im$ with all images $dim(im) = (17, 60, 125, 3)$
where 17 is $n$ the number of images and (60,125,3) are each image dimensions

```{r}
im=array(rep(0,length(files)*height/scale*width/scale*3), dim=c(length(files), width/scale, height/scale, 3))

for (i in 1:n){
  im[i,,,]=resizeImage(readImage(paste0(path, files[i])),width=width/scale, height=height/scale)
  }
```

Plot first 9 images

```{r}
par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))

for (i in 1:9){
  plot_jpeg(writeJPEG(im[i,,,]))
}
```

To begin PCA each image is flatten into a matrix where each column is one image
$mshoes$ will be the matrix of  $dim(im) = 22500 x 17$

```{r}
flat=matrix(0, n, prod(dim(im)[2:4]))
for (i in 1:n) {
  newim <- readImage(paste0(path, files[i]))
  r=as.vector(im[i,,,1]) 
  g=as.vector(im[i,,,2])
  b=as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
mshoes=t(flat)
```


The principal components in PCA are the eigen vectors of the covariance matrix

We compute the covariance matrix by first standardizing each image by $(X - \mu) / \sigma$
Once standardized each image will have a 
$\mu=0\\\sigma=1$

With the standarized values we compute the covariance matrix to get the linear combination 
between each shoe, the shape of the covariance matrix is 17x17

```{r}
scaled=scale(mshoes, center = TRUE, scale = TRUE)
means=attr(scaled, "scaled:center") #saving for classification
std=attr(scaled, "scaled:scale") 
cshoes=cov(scaled)
```


With the covariance matrix the eigen values are corresponding eigen vectors can be computed

```{r}
edf=eigen(cshoes)
```

The eigen values tell you how much variance can be captured with the $ith$ eigen values

```{r}
lambda_dist<-as.data.frame(t(cumsum(edf$values) / sum(edf$values)))
lambda_dist
```

It looks like the first 3 eigen values can be used to represent more than 80% of the data with 6 we are at 90%
and by 11 it's 95% of the pixel variance.

Now to visualize what an image looks like with only 80%

```{r}
scaling=diag(edf$values[1:3]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes=scaled%*%edf$vectors[,1:3]%*%scaling
imageShow(array(eigenshoes[,1], c(width/scale,height/scale,3)))
```