---
title: "Homework7 - DATA 605"
author: "Kenan Sooklall"
date: "3/10/2021"
output:
  pdf_document: default
  html_document: default
---

### Question 1
1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

Each Xi has k possibilities [1,2,3....k] for i in K therefore for all possibilities of X's = k^n
The number of way to get Y= 1 -> k^n ways to get 1 and (k-1)^n to not get 1
So 

$P(X=1) = k^n * (k-1)^n / k^n \\$
$\rightarrow P(X=2) = (k-2+1)^n * (k-2)^n / k^n \\$
$\rightarrow P(X=3) = (k-3+1)^n * (k-3)^n / k^n \\$

$\rightarrow P(X=j) = (k-j+1)^n * (k-j)^n / k^n$

### Question 2
2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

```{r}
p = 1/10
```

#### Geometric
a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

$ P(X = k) = (1-p) ^ {k - 1}*p$

P(X>8) = 1 - P(X<=8)

Expected Value = 1/ 0.1 = 10

Standard deviation = sqrt((1-p)/p^2) = 9.48

```{r}
1 - pgeom(8-1,p)
(1-p)^(8-1)*p
```

#### Exponential

b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.

$P (X >= k) = e^{-k/\mu}$

Expected Value = 10

Standard deviation = 10

P(X>=8)

```{r}
1 - pexp(8, p)
exp(-8/10)
```

#### Binomial

c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

$P (X > k) = p ^ k * ( 1-p ) ^{n-k}$

Expected Value = 8 * 0.1 = 0.8

Standard deviation = sqrt(8 * 0.1 * 0.9) = 0.85

```{r}
k=0
n=8
pbinom(0, size=n, prob=p)
0.1^0 * (0.9) ^ (8-0)
```

#### Poisson

d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson.

$P(X = 8) = \lambda ^ x e^{-\lambda} / (x!)$

Expected Value = 8 * 0.1 = 0.8

Standard deviation = sqrt(0.8) = 0.89

```{r}
ppois(0, 0.8)
0.8 ^ 0 * exp(-0.8) / (factorial(0))
```
