---
title: "Homework3"
author: "Kenan Sooklall"
date: "2/6/2021"
output:
  html_document: default
  pdf_document: default
---

### Problem set 1
(1) What is the rank of the matrix A?
\[
A=
  \begin{bmatrix}
    1 & 2 & 3 & 4 \\
    -1 & 0 & 1 & 3 \\
    0 & 1 & -2 & 1 \\
    5 & 4 & -2 & -3
  \end{bmatrix}
\]

```{r}
library(Matrix)
m <- matrix(c(c(1,-1,0,5), c(2,0,1,4), c(3,1,-2,-2), c(4,3,1,-3)), nrow=4)
rankMatrix(m)
```
The rank of matrix $A$ is 4, each row is linearly independent

(2) Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?

The maximum rank a matrix of the shape mxn where m>n can be is m, the max(m, 1).
The minimum rank a non-zero matrix can have is 1 where all rows are a linear combination of one row.

(3) What is the rank of matrix B?
\[
B = \begin{bmatrix}
    1 & 2 & 1 \\
    3 & 6 & 3 \\
    2 & 4 & 2 \\
    \end{bmatrix}
\]

```{r}
m <- matrix(c(c(1,3,2), c(2,6,4), c(1,3,2)), nrow=3)
rankMatrix(m)
```
Rows 2 and 3 are a linear combination of row 1, 

$row 2 =  row1 * 3 \\$ 
$row 3 = row1 * 2 \\$

Therefore the rank of matrix B = 1

### Problem set 2

Compute the eigenvalues and eigenvectors of matrix A. You'll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.

\[
A = \begin{bmatrix}
    1 & 2 & 1 \\
    0 & 4 & 5 \\
    0 & 0 & 6 \\
    \end{bmatrix}
\]

Please show your work using an R-markdown document. Please name your assignment
submission with your first initial and last name.

#### eigenvalues


Starting with $det(A -	\lambda * I) = 0$

Then

\[
X = \begin{bmatrix}
    \lambda & 0 & 0 \\
    0 & \lambda & 0 \\
    0 & 0 & \lambda \\
    \end{bmatrix}
    -
    \begin{bmatrix} 
	1 & 2 & 3 \\ 
	0 & 4 & 5 \\ 
	0 & 0 & 6 \\ 
	\end{bmatrix}
	=
	\begin{bmatrix}
    \lambda - 1 & -2 & -3 \\
    0 & \lambda - 4 & -5 \\
    0 & 0 & \lambda - 6 \\
    \end{bmatrix}
\]

Next find the determinant

$det(X) = 0$

$(\lambda - 1)(\lambda - 4)(\lambda - 6) - (-2)(0) - 3(0) = (\lambda - 1)(\lambda^2 - 10\lambda + 24) = \lambda^3 - 11\lambda^2 + 3\lambda - 24 = 0$

Solving that cubic equation and assuming integer solutions. Factors of 24 are [1, 2, 3, 4, 6, 8, 12, 24], 12 and 24 will blow up with the cube so will check those last

$1: 1^3 - 11 * 1^2 + 34 * 1 - 24 = -10 + 34 - 24 = 0 \quad\quad \checkmark \\$
$2: 2^3 - 11 * 2^2 + 34 * 2 - 24 = 8 + 24 - 24 = 8   \quad\quad\quad X \\$
$3: 3^3 - 11 * 3^2 + 34 * 3 - 24 = 27 + 3 - 24 = 6   \quad\quad\quad X \\$
$4: 4^3 - 11 * 4^2 + 34 * 4 - 24 = 64 - 40 - 24 = 0 \quad\quad\quad \checkmark \\$
$6: 6^3 - 11 * 6^2 + 34 * 6 - 24 = 216 - 192 - 24 = 0 \quad\quad \checkmark$

3 eigenvalues

$\lambda_1 = 1 \\$
$\lambda_2 = 4 \\$
$\lambda_3 = 6 \\$ 


\newpage
#### eigenvectors

For $\lambda_1 = 1 \\$

$A -\lambda I$ = 
\[
\begin{bmatrix} 
	0 & 2 & 3 \\ 
	0 & 3 & 5 \\ 
	0 & 0 & 5 \\ 
	\end{bmatrix}
	\begin{bmatrix} 
	x \\y \\ z
	\end{bmatrix}
	=0
\]

Simplifying

$2y + 3z=0 \\$
$3y + 5z=0 \\$
$5z=0 \\$

$z = 0 \rightarrow y=0 \rightarrow x=x \\$

Getting the vector

\[
x\begin{bmatrix} 
	1 \\0 \\ 0
	\end{bmatrix}
\]

- $\lambda_1 = 1 \quad \overrightarrow V_1 = (1, 0, 0)$

\newpage
For $\lambda_1 = 4$

$A -\lambda I$ = 

\[
\begin{bmatrix} 
	-3 & 2 & 3 \\ 
	0 & 0 & 5 \\ 
	0 & 0 & 2 \\ 
	\end{bmatrix}
	\begin{bmatrix} 
	x \\y \\ z
	\end{bmatrix}
	=0
\]

Simplifying
$-3x+2y + 3z=0 \\$ 
$5z=0 \\$
$2z=0 \\$

$z=0 \rightarrow 2y=3x \rightarrow x=2, y=3$

Getting the vector
\[
\begin{bmatrix} 
	2 \\3 \\0
	\end{bmatrix}
\]

- $\lambda_1 = 4 \quad \overrightarrow V_2 = (2, 3, 0)$

\newpage
For $\lambda_1 = 6$

$A -\lambda I$ = 

\[
\begin{bmatrix} 
	-5 & 2 & 3 \\ 
	0 & -2 & 5 \\ 
	0 & 0 & 0 \\ 
	\end{bmatrix}
	\rightarrow[\text{RREF}]
	\begin{bmatrix} 
	-5 & 0 & 3 \\ 
	0 & -2 & 5 \\ 
	0 & 0 & 0 \\ 
	\end{bmatrix}
	\begin{bmatrix} 
	x \\y \\ z
	\end{bmatrix}
	=0
\]

Simplifying
$-5xy + 8z=0 \\$
$-2y + 5z=0 \\$

$8z=5x \\$
$5z=2y \\$

$x=8, z=5, y=12.5 \\$

Multiply by $2$ for integer values

\[
	\begin{bmatrix} 
	16 \\25 \\10
	\end{bmatrix}
\]

- $\lambda=6 \quad \overrightarrow V_3 = (16, 25, 10)$



