---
title: "Lab 1: Intro to R"
author: "Kenan Sooklall"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

```{r trend-girls}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point() +  geom_smooth()
```

Is there an apparent trend in the number of girls baptized over the years? How would you describe it? (To ensure that your lab report is comprehensive, be sure to include the code needed to make the plot as well as your written interpretation.)

- The plot is showing an increase in the number of girls being baptized from 1629 to 1710.

### Exercise 3

Now, generate a plot of the proportion of boys born over time. What do you see?

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line() + 
  geom_smooth()
```

- It looks like the proportion of boys initially went up, then dropped and then slowly rising.

### Exercise 4

What years are included in this data set? What are the dimensions of the data frame? What are the variable (column) names?

```{r dim-present}
glimpse(present)
unique(present$year)
NROW(present); NCOL(present)
```

- The data set is 63 rows and 3 columns (year, boys, girls) which includes years 1940 to 2002.


### Exercise 5

How do these counts compare to Arbuthnot’s? Are they of a similar magnitude?
```{r count-compare}
present <- present %>%
  mutate(total = boys + girls)
head(present)
```

- The values in present are much bigger than arbuthnot by 3 orders of magnitude

### Exercise 6

Make a plot that displays the proportion of boys born over time. What do you see? Does Arbuthnot’s observation about boys being born in greater proportion than girls hold up in the U.S.? Include the plot in your response. Hint: You should be able to reuse your code from Exercise 3 above, just replace the dataframe name.

```{r plot-prop-boys-present}
present <- present %>%
  mutate(total = boys + girls)
present <- present %>%
  mutate(boy_to_girl_ratio = boys / girls)
present <- present %>%
  mutate(boy_ratio = boys / total)

ggplot(data = present, aes(x = year, y = boy_ratio)) + 
  geom_line()
```

- The graph is very different from Arbuthnot's observation, the boy_ratio is decreasing with time.

### Exercise 7

In what year did we see the most total number of births in the U.S.? Hint: First calculate the totals and save it as a new variable. Then, sort your dataset in descending order based on the total column. You can do this interactively in the data viewer by clicking on the arrows next to the variable names. To include the sorted result in your report you will need to use two new functions: arrange (for sorting). We can arrange the data in a descending order with another function: desc (for descending order). The sample code is provided below.

```{r find-max-total}
# Insert code for Exercise 7 here
present[which.max(present$total), ]
```

- It looks like year 1961 had the most total number of births in the U.S at 4268326.
