---
title: "Homework 8 DATA-605"
author: "Kenan Sooklall"
date: "3/15/2021"
output:
  html_document: default
  pdf_document: default
---

#### Exercise 11 
A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out? (See Exercise 10.)

```{r}
n=100
lifetime=1000
lifetime/n
```

#### Exercise 14 
Assume that X_1 and X_2 are independent random variables, each having an
exponential density with parameter $\lambda$. 
Show that $Z = X 1 − X 2$ has density

$f_Z(z) = (1/2)λe^{−λ|z|}$

Expanding the solution

$$f_Z(z) = λ/2e^{−λ|z|} \; \begin{cases} 
      λ/2e^{−λz} & z >= 0 \\
      λ/2e^{λz} & z < 0 \\
   \end{cases}$$
   
let $X = X_1 \& Y = X_2 \rightarrow Z=X-Y \rightarrow Y = X - Z$

For $z >= 0 \rightarrow x \in (z , \infty)$ 

$$ \int_{z}^{\infty}\lambda e ^{-\lambda x} \lambda e ^ {-\lambda(x-z)}dx = \lambda^2\int_{z}^{\infty} e ^{\lambda(z - 2x)} dx $$
Using u substitution 

$u = \lambda * (z - 2x) \rightarrow du=-2\lambda dx$

Simplify
$$(-\frac{\lambda}{2}e^ {z - 2x}|_z^{\infty}) = \frac{\lambda}{2}e^{-\lambda z}$$
For $z < 0 \rightarrow x \in (0 , \infty)$

$$(-\frac{\lambda}{2}e^ {z - 2x}|_0^{\infty}) = \frac{\lambda}{2}e^{\lambda z}$$

#### Exercise 1
Let X be a continuous random variable with mean $\mu = 10$ and variance
$σ^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the
following probabilities

$\sigma = 5.77$

(a) $P (|X − 10| ≥ 2) \rightarrow 2 = k\sigma \rightarrow k=0.346 \rightarrow 1/k^2 = 0.12$

(b) P (|X − 10| ≥ 5)

```{r}
std = sqrt(100/3)
k = 5/std
1/k^2
```

(c) P (|X − 10| ≥ 9).

```{r}
k = 9/std
1/k^2
```

(d) P (|X − 10| ≥ 20).

```{r}
k = 20/std
1/k^2
```