---
title: "Homework_3"
author: "Kenan Sooklall"
date: "2/10/2021"
output:
  pdf_document: default
  html_document: default
---

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS" 
```{r}
library(tidyverse)
df <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')
df %>% filter(across(Major, ~ grepl('DATA|STATISTICS', .)))
```

#2 Write code that transforms the data below:

```{r}
df <- data.frame(x1=c("bell pepper", "bilberry", "blackberry", "blood orange"), x2=c("blueberry", "cantaloupe", "chili pepper", "cloudberry"), 
                x3=c("elderberry", "lime", "lychee", "mulberry"))
```
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

```{r}
vect <- as.vector(as.matrix(df[,c("x1", "x2", "x3")]))
vect
```
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

Regex checked on https://regexr.com/ and https://regex101.com/
  
(.)\1\1

Matches a text that is repeated 3 times like 
Will match: (111)1(222)2)(333)(aaa)z(bbb)d(ccc)

(.)(.)\\2\\1

Matches any character for the first two steps then the number 2 and 1 separated by a backslash
Will match: ab\2\1 AND bb\2\1
Will not match: xy\3\1

(..)\1

Two capture groups, matches the first pair then the (\1) will match those pairs. A total of 4 characters will match
Will match: (2222)(3333)
Will not match: aaac

(.).\\1.\\1

Matches any first two character then (\1) exactly then any other character then a (\1)
Will match: qb\1f\1 ab\1c\1
Will not match: qb\1f\2 ab\4d\2

(.)(.)(.).*\\3\\2\\1

Matches any first 3 character then an infinate amount of characters followed by (\3)(\2)(\1) exactly
Will match: abcwgasdhrhrana\3\2\1
Will not match: dgasdhgaheghjlhlasg\4\2\1

#4 Construct regular expressions to match words that:

Start and end with the same character.

"(^.).*\1$"

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

"(..).*\1"

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

([a-z]).*\1.*\1

